Transient Corotating Clumps Around Adolescent Low-Mass Stars From Four Years of TESS

### 2023-09-07 paper/ms.tex
update ms

% 250 word max!
\begin{abstract}
  %
  Complex periodic variables (CPVs) are stars that exhibit sharp and
  periodic optical light curves.  Previous studies have indicated that
  these stars are typically disk-free pre-main-sequence M dwarfs with
  rotation periods ranging from 0.2 to 2 days.  To advance our
  understanding of these enigmatic objects, we conducted a blind
  search using TESS 2-minute data of \nstarssearched\ K and M dwarfs
  with $T$$<$16 and $d$$<$150\,pc.  We found \ngoods\ high-quality
  CPVs, and subsequently determined that most are members
  of stellar associations.  Among the new discoveries are the
  brightest ($T$$\approx$9.5), closest ($d$$\approx$20\,pc), and
  oldest ($\approx$200\,Myr) CPVs known.  One exceptional object, LP
  12-502, exhibited up to eight flux dips per cycle. Some of these
  dips coexisted with slightly different periods, and the
  shortest-duration LP 12-502 dips precisely matched the expected
  timescale for transiting small bodies at the corotation radius.
  Broadly, our search confirms that CPVs are mostly young
  ($\lesssim$150\,Myr) and low-mass ($\lesssim$0.4\,$M_\odot$).  The
  flux dips characteristic of the class have lifetimes of $\approx$100
  cycles, although stellar flares seem to induce sudden dip collapse
  once every few months.  The most plausible explanation for these
  phenomena remains corotating concentrations of gas or dust.  The gas
  or dust is probably entrained by the star's magnetic field, and the
  sharp features could result from a multipolar field topology, a
  hypothesis supported by correspondences between the light curves of
  CPVs and of rapidly rotating B stars known to have multipolar magnetic fields.
  %
\end{abstract}

@@

{\it Multiple periodic TESS signals}---During our visual analysis of
the TESS light curves and PDM periodograms, we flagged sources with
beating light curves, and with PDM periodograms that showed multiple
periods.  In such cases, we attempted to disentangle the two signals
by subtracting the mean CPV signal over each sector, and we then
repeated the phase-dispersion minimization analysis.  The resulting
secondary periods, $P_{\rm sec}$, are listed in
Table~\ref{tab:thetable}; we required these to be at least 5\%
different from the primary period, and for the majority of secondary
signals the morphologies corresponded to starspot modulation.  This
process yielded \ngoodmultperiodflag/\ngoods\ high-quality CPVs with
secondary periods; \nmaybemultperiodflag/\nmaybes\ of the ambiguous
sample met the same criterion.  Of the \ngoodhighruwe\ good CPVs with
${\rm RUWE}_{\rm DR3}> 2$, \ngoodruweandmultperiod\ also showed
secondary periods in the TESS light curves.  Considering the weaker
threshold of ${\rm RUWE}_{\rm DR3}> 1.4$,
\ngoodweakruweandmultperiod/\ngoodweakruwe\ such CPVs showed secondary
TESS periods.  These results broadly suggest that the secondary
periods are associated with physically-bound binary companions.

Table~\ref{tab:thetable} summarizes each of the sources of
binarity information into a single bitwise column.  
We present detailed results concerning binarity in
Section~\ref{subsec:resultsbinarity}, and a summary discussion in
Section~\ref{subsec:discbinary}.

@@

To date, the only stars reported to show the CPV phenomenon are M
dwarfs, with typical stellar masses $\lesssim$0.3\,$M_\odot$
\citep{2017AJ....153..152S,2022AJ....163..144G}.  However the two most
massive CPVs in our sample, TIC~405754448 and TIC~405910546, were
assigned masses of $\approx$0.82\,$M_\odot$ and
$\approx$0.60\,$M_\odot$ respectively.  The next-highest masses in our
sample are $\approx$0.40\,$M_\odot$.

These stars' locations in color--absolute magnitude diagrams and their
probable membership in Lower Centaurus Crux both support the
conclusion that they have relatively high masses.  However in detail,
both are subject to ambiguities in interpretation.  The TIC~405910546
light curve has a unique shape, suggestive of an eclipsing binary.
Independently, TIC~405910546 was one of only \nrvscatterflag\ CPVs
flagged with a Gaia DR3 radial velocity scatter exceeding 20\,\kms.
Combined, these factors suggest that TIC~405910546 could be a
pre-main-sequence eclipsing binary; it should be studied further to
clarify this classification.

@@

	during its state-switches (e.g.\ Figure~\ref{fig:lp}, cycles 1233-1264)
	imply that the flux dips are additive and independent.

@@

      moving group (TIC~294328887; $\approx$200\,Myr).  The
      lack of detected CPVs in the Hyades and Praesepe suggests that
      the lifetime of the phenomenon is limited to the first few
      hundred million years.

@@

    \item Most CPVs are M dwarfs with masses
      0.1-0.5\,$M_\odot$.  Two sources, TIC~405754448 and
      TIC~405910546, have masses that appear to exceed the M dwarf
      limit.  Both are potentially binaries, and this may
      confuse our ability to accurate identify the source of the CPV
      signal (Section~\ref{subsec:massive}).  We encourage additional

>>>>

% 250 word max!
\begin{abstract}
  %
  Complex periodic variables (CPVs) are stars that exhibit 
  highly structured and periodic optical light curves.  Previous
  studies have indicated that these stars are typically disk-free
  pre-main-sequence M dwarfs with rotation periods ranging from 0.2 to
  2 days.  To advance our understanding of these enigmatic objects, we
  conducted a blind search using TESS 2-minute data of
  \nstarssearched\ K and M dwarfs with $T$$<$16 and $d$$<$150\,pc.  We
  found \ngoods\ high-quality CPVs, and subsequently determined that
  most are members of stellar associations.  Among the new discoveries
  are the brightest ($T$$\approx$9.5), closest ($d$$\approx$20\,pc),
  and oldest ($\approx$200\,Myr) CPVs known.  One exceptional object,
  LP 12-502, exhibited up to eight flux dips per cycle. Some of these
  dips coexisted with slightly different periods, and the
  shortest-duration dips precisely matched the expected
  timescale for transiting small bodies at the corotation radius.
  Broadly, our search confirms that CPVs are mostly young
  ($\lesssim$150\,Myr) and low-mass ($\lesssim$0.4\,$M_\odot$).  The
  flux dips characteristic of the class have lifetimes of $\approx$100
  cycles, although stellar flares seem to induce sudden dip collapse
  once every few months.  The most plausible explanation for these
  phenomena remains corotating concentrations of gas or dust.  The gas
  or dust is probably entrained by the star's magnetic field, and the
  sharp features could result from a multipolar field topology, a
  hypothesis supported by correspondences between the light curves of
  CPVs and of rapidly rotating B stars known to have multipolar
  magnetic fields.
  %
\end{abstract}

@@

{\it Multiple periodic TESS signals}---During our visual analysis of
the TESS light curves and PDM periodograms, we flagged sources with
beating light curves, and with PDM periodograms that showed multiple
periods.  For such cases, we then
subtracted the mean CPV signal over each sector, and 
repeated the phase-dispersion minimization analysis.  The resulting
secondary periods, $P_{\rm sec}$, are listed in
Table~\ref{tab:thetable}; we required these to be at least 5\%
different from the primary period.  The majority of secondary
signals showed morphologies corresponding to starspot modulation.  This
process yielded \ngoodmultperiodflag/\ngoods\ high-quality CPVs with
secondary periods; \nmaybemultperiodflag/\nmaybes\ of the ambiguous
sample met the same criterion.  Of the \ngoodhighruwe\ good CPVs with
${\rm RUWE}_{\rm DR3}> 2$, \ngoodruweandmultperiod\ also showed
secondary periods in the TESS light curves.  Considering the weaker
threshold of ${\rm RUWE}_{\rm DR3}> 1.4$,
\ngoodweakruweandmultperiod/\ngoodweakruwe\ such CPVs showed secondary
TESS periods.  The latter results strongly suggest that the secondary
periods are associated with bound binary companions.

Table~\ref{tab:thetable} summarizes each of the sources of
binarity information into a single bitwise column.  
We describe detailed results concerning binarity in
Section~\ref{subsec:resultsbinarity}, and summarize those results
Section~\ref{subsec:discbinary}.

@@

To date, the only stars reported to show the CPV phenomenon are M
dwarfs, with typical stellar masses $\lesssim$0.3\,$M_\odot$
\citep{2017AJ....153..152S,2022AJ....163..144G}.  However the two most
massive CPVs in our sample, TIC~405754448 and TIC~405910546, were
assigned masses of $\approx$0.82\,$M_\odot$ and
$\approx$0.60\,$M_\odot$ respectively.  The next-highest mass in our
sample belongs to TIC~59836633 ($\approx$0.48\,$M_\odot$), with all
remaining CPVs having masses $\lesssim$0.40\,$M_\odot$.

The locations of TIC~405754448 and TIC~405910546 in color--absolute
magnitude diagrams, combined with their probable membership in Lower
Centaurus Crux, support the conclusion that these stars have
relatively high masses.  However in detail, both objects are subject
to ambiguities in interpretation.  The TIC~405910546 light curve has a
unique shape, suggestive of an eclipsing binary.  Independently,
TIC~405910546 was one of only \nrvscatterflag\ CPVs flagged with a
Gaia DR3 radial velocity scatter exceeding 20\,\kms.  Combined, these
factors suggest that TIC~405910546 could be a pre-main-sequence
eclipsing binary; it should be studied further to clarify this
classification.

@@

      during its state-switches (e.g.\ Figure~\ref{fig:lp}, cycles
      1233-1264) imply that the flux dips are additive and
      independent.

@@

      moving group (TIC~294328887; $\approx$200\,Myr).  The lack of
      detected CPVs in the Hyades and Praesepe suggests that the
      lifetime of the phenomenon is limited to the first few hundred
      million years.

@@

    \item Most CPVs are M dwarfs with masses 0.1-0.4\,$M_\odot$.
    Two sources, TIC~405754448 and
      TIC~405910546, have masses that appear to exceed 0.5\,$M_\odot$.
      Both are potentially binaries, and this may confuse our ability
      to accurate identify the source of the CPV signal
      (Section~\ref{subsec:massive}).  We encourage additional
